The `Maltreatment and Abuse Chronology of Exposure' (MACE) scale is essentially unidimensional

### 2022-07-26 manuscript/main.tex
backing up before refactorization
manuscript updates

@@

\title{Bifactor models of the `Maltreatment and Abuse Chronology of Exposure' (MACE) Scale}
\shorttitle{Bifactor models of MACE}
\author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}

@@

\begin{document}
\maketitle

@@

% "Happy families are all alike; every unhappy family is unhappy in its own way." – Leo Tolstoy.\break

Childhood maltreatment is a common experience for children across the world \citep{stoltenborgh2015prevalence}. It leads to an array of poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, adverse effects on cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}. Furthermore, childhood maltreatment predicts worse treatment outcomes across a variety of psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}. As such, assessment of childhood maltreatment in a reliable and valid manner is important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design interventions to prevent maltreatment in the first place.

Many measures of childhood abuse and maltreatment have been developed \citep{saini2019systematic}. One recently introduced measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages. The MACE is a 52-item self-report questionnaire of childhood maltreatment that assays 10 unique types of maltreatment. Six of these (i.e. parental physical maltreatment, parental non-verbal emotional abuse, parental verbal abuse, sexual abuse, physical neglect, emotional neglect) share overlap with other common measures of abuse (e.g. CTQ \citep{bernstein1998childhood} ); the remaining four solicit additional information on peer abuse, and witnessing violence at home. A number of previous scales were reviewed to select candidate items for the MACE. Item response theory and Rasch model were then used to for final selection of items for each 10 subscales. The MACE also measures detailed information on timing of exposure, which is critical for investigating how the onset of abuse shapes trajectories of development. Designed under the cumulative risk framework \citep{evans2013cumulative}, the MACE provides an overall severity score and multiplicity score (number of types of maltreatment experienced) with good test-retest reliability, convergent validity, and external validity.

More recently, there have been calls to move away from cumulative risk approach \citep{evans2013cumulative} --- which focuses on the number of adverse experiences, but does not distinguish between distinct types of environmental experience --- towards a dimensional approach, which emphasizes how specific types of maltreatment uniquely shape psycho-biological developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. As such, research groups have begun to use the MACE in other ways. Some groups have studied the unique contributions of all 10 subscales to psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Following the threat-deprivation dimensional model \citep{mclaughlin2014childhood}, others have made composite indices of threat and deprivation using the MACE items measuring abuse and neglect, respectively \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}.

However, a critical challenge in measuring childhood maltreatment is the high co-occurrence of multiple types of abuse and neglect experiences; that is, children experiencing maltreatment are likely to be exposed to two or more types of maltreatment \citep{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}. For example, virtually all severely maltreated children report emotional abuse \citep{mcgee1995measurement}. As a consequence, isolating the unique contribution of a specific type of maltreatment (e.g. emotional neglect) to the risk of developing some outcome (e.g. depression) is difficult insofar that any one type of adversity is likely confounded with general maltreatment. Concretely, this means that researchers attempting to study the effects of particular types of maltreatment using subscale scores (e.g. emotional neglect from the MACE scale) should take pains to ensure that those measures largely reflect a construct unique from general maltreatment, as represented by the total score.

Bifactor models, a group of confirmatory factor analyses, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. Bifactor models specify that the covariance among the items can be accounted for by a latent general factor, reflecting common variance among all items, and a set of latent specific factors, reflecting additional common variance for subsets of items \citep{Reise2012-ql}. Crucially bifactor models can be used to calculate model-based reliability statistics, e.g. the omega-hierarchical ($\omega_h$) and omega-subscale ($\omega_s$) indices, which respectively measure the proportion of variance in the total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}. Applied to MACE scale, these indices can guide researchers as to whether subscale scores for particular types of maltreatment are interpretable as such, or if they instead largely reflect general maltreatment. 
% summary of paper paragraph tk

% The bifactor model hypothesizes a general factor, onto which all items load, and a series of orthogonal (uncorrelated) skill-specific grouping factors
\section{Methods}

@@

The second sample of participants (N=687) were recruited from the Prolific Academic platform (\url{https://www.prolific.co}) to participate in an online experiment in August -- October, 2021. Of these, N=582 were retained for analysis (see Exclusion Criteria below). Participants were eligible for participation if they were 18 years or older and currently resided in the United States or Canada. Participants received monetary compensation for their time (rate: \$12 USD/hr). This study was approved by the Institutional Review Board of the University of Ottawa, and all participants provided informed consent. This sample will hereafter be referred to as the \textit{replication} sample.

@@

The fit indices for each model and sample are summarized in Table \ref{table:diagnostics}. According to the traditional indices, we found all four models provided an acceptable fit to the response data. Across samples, all the models had RMSEA values < 0.06 and the majority exhibited CFI and TLI values > 0.95. The exception were models 1 and 3 in the original sample, which exhibited acceptable fits (CFI and TFI values > 0.90). In general, model fits were better for the replication sample data than for the original sample. Furthermore none of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure exceeded the critical value, indicating that all models were sufficiently able to replicate the sum score distribution for each sample.

Next we inspected the $Q_3$ indices for evidence of local dependence. For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value indicating the presence of local dependence. To understand the extent of the local dependence, we visualized the $Q_3$ values for item pairs exceeding the critical value for each model and sample (Figure S3). For both samples the unidimensional model exhibited many item pairs with local dependence, predominantly among the peer victimization and reverse-scored items, suggesting possible underfactorization. By comparison, the remaining models exhibited far fewer locally dependent item pairs, the majority of which were easy to understand. For example, the large $Q_3$ value for item 17 (``Parents made inappropriate sexual comments or suggestions to your sibling or stepsibling'') and item 18 (``Parents touched or fondled your sibling or stepsibling in a sexual way'') is easily understood as surface local dependence. Regardless, the estimated item discrimination parameters for all models and samples seldom exceeded the usual range (i.e. $\alpha \leq 4$) suggesting that whatever local dependence was present is exerting minimal bias on parameter estimates \citep{edwards2018diagnostic}. Thus, we conclude that all of the models provide at least acceptable fits to the response data from both samples.

The results of the model comparison are also summarized in Table \ref{table:diagnostics}. In both the original and replication datasets, models 2 and 4 exhibited better fit to the data than models 1 and 3. Between models 2 and 4, the former provided a numerically better fit in the original sample ($\Delta$ LOO = 7.9, se = 56.3) but a worse fit in the replication sample ($\Delta$ LOO = -24.0, se = 52.0). The differences in LOO-CV values are both within four standard errors of the mean, however, indicating only weak predictive improvements \citep{vehtari2022cv}. Thus, the model comparison suggests that models 2 and 4 yield approximately equal fits to the data, which is noteworthy insofar that model 4 assumes a much simpler and parsimonious factor structure of the MACE scale.

The standardized factor loadings from the unidimensional model fit to the original and replication datasets are are summarized in the first column of Figure \ref{fig:loadings_original} and Figure \ref{fig:loadings_online} respectively.

\subsection{Model 1: Unidimensional model}

To begin we fit a unidimensional model to the MACE response data, in which all items load onto a single general maltreatment factor (Figure 1a). In the original dataset, the item discrimination parameters were between 0.506 and and 2.944, with an average value of 1.357 (95\% HDI = 1.295 -- 1.414). For the replication dataset, item discrimination ranged between 0.657 and 2.407 with an average value of  1.349 (95\% HDI = 1.272 -- 1.422). Across datasets, all but one item exhibited medium ($\alpha \in$ 0.65 -- 1.34) to high ($\alpha \in$ 1.35 -- 1.69) levels of discrimination \citep{baker2017basics}. Differences in item discrimination estimates between datasets were marginal (RMSE = 0.038, 95\% HDI = 0.014 -- 0.068), indicating consistency of discrimination across samples.

In the original dataset, factor loadings ranged between 0.279 and 0.864, with an average value of 0.597 (95\% HDI = 0.579 -- 0.614). Similarly, in the replication dataset, factor loadings ranged between 0.359 and 0.814 with an average value of 0.600 (95\% HDI = 0.578 -- 0.621). With the exception of item 42, all items across datasets exhibited consistent and moderate-to-large loadings on the unitary factor. In other words, all but one item appear to measure, at least in part, general maltreatment.

\subsection{Model 2: Bifactor model with 10 specific factors}

The standardized factor loadings from the fit of the bifactor model to the original and replication datasets are are summarized in the second column of Figure \ref{fig:loadings_original} and Figure \ref{fig:loadings_online} respectively. For both datasets, the factor loadings on the general factor largely resembled those from the unidimensional models. The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset. The relative parameter bias\footnote{The relative parameter bias is defined as the difference between an item’s loading in the unidimensional solution and its general factor  loading in the bifactor, divided by the general factor loading in the bifactor model.} was 2.1\% for the original dataset and 2.7\% in the replication dataset, indicating little bias in the factor loadings in a unidimensional solution due to multidimensionality.

In contrast loadings on the specific factors were predominantly small-to-moderate in magnitude, and seldom exceeded loadings on the general factor. In the original dataset, specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341). For the replication dataset, the range was 0.043 to 0.698 with an average value of 0.343 (95\% HDI = 0.315 -- 0.369). Only items from the peer victimization subscales (i.e. PeerVA, PeerPA) had specific factor loadings of roughly equal or greater magnitude than those of general factor.

Next we inspected the reliability of the MACE total score. The $\omega$ values indicated the total score had excellent reliability in both the original ($\omega$ = 0.963) and replication ($\omega$ = 0.965) datasets. Across datasets, the omega hierarchical values was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in total scores are attributable to general maltreatment. Moreover, the explained common variance values were 0.716 for the original dataset and 0.697 for the replication dataset. This indicates the presence of a strong general maltreatment factor in the dataset, and suggests MACE response data is largely undimensional.

Next we inspected the reliability of the MACE subscale scores. The model-based reliability indices were variable in both the original (mean $\omega$ = 0.750, range = 0.590 -- 0.909) and replication (mean $\omega$ = 0.756, range = 0.595 -- 0.893) datasets, though most subscale scores exhibited adequate reliability. Crucially, the omega subscale were typically small in both the original (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565) replication datasets. This indicates that most subscale scores contain little reliable variance apart from the general factor; that is, the majority of MACE subscales scores cannot be meaningfully interpreted as such. The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores were independent of general maltreatment (original = 70.7\%; replication = 66.2\%). These findings were echoed in the H index values, which were unacceptably small for all subscales across both datasets (with an exception again for the peer verbal abuse subscale).

\subsection{Model 3: Bifactor S-1 model with one specific factor (neglect)}

As for the full bifactor model, the factor loadings on the threat-deprivation bifactor S-1 model resembled those from the unidimensional models. The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset. The standardized factor loadings for the specific factor (i.e. deprivation) were on average [stats here]. The loadings, however, exhibited a striking pattern where only the reverse-scored items exhibited moderate loadings; the the remaining four items exhibited negligible loadings. That is, the "deprivation" factor appears to reflect a methods factor instead.

Though the subscale reliability was adequate (original: $\omega$ = 0.877; replication: $\omega$ = 0.921), the omega subscale values indicated that neglect scores contained less than half reliable variance unique from the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375). Not only is the validity of the deprivation subscale questionable, but so too is its reliability.

\subsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}

Again the standardized factor loadings on general factor of the parsimonious bifactor S-1 model resembled those from the unidimensional models. The average magnitude of the factor loadings was 0.575 (95\% HDI = 0.556 -- 0.593) for the original dataset and 0.580 (95\% HDI = 0.560 -- 0.601) for the replication dataset. Notably, the factor loadings of the peer victimization and methods factor were also moderate in magnitude and on average larger than those observed on the general factor, lending credence for their independence.

In contrast, the peer victimization factor of the second bifactor S-1 model (Model 4) exhibited good subscore reliability in both datasets original: $\omega$ = 0.889; replication: $\omega$ = 0.868) and adequate omega subscale reliability (original: $\omega_h$ = 0.569; replication: $\omega_h$ = 0.493). In other words, more half of its reliable variance was unique from the general factor (original = 0.640; replication = 0.568).

@@

The standardized factor loadings for the two- and three-factor solutions generated by the cf-quartimax rotation are summarized in Figure \ref{fig:efa}. (The corresponding factor loadings yielded by the geomin rotation were largely similar and are presented Figure S1.) In the two-factor solutions, two unique factor structures emerged by dataset. In the original dataset, there was a dominant household maltreatment factor and a secondary peer victimization factor. In contrast, the replication dataset exhibited a primary factor comprised of household and peer maltreatment, and a secondary reverse-scoring methods factor.

In the three-factor solutions, the recovered factor structures were largely similar across datasets. There was a household maltreatment factor, a peer victimization factor, and a reverse-scoring methods factor. Of note, the sexual abuse items seemed to load more strongly on the peer victimization factor in the replication dataset than in the original dataset, warranting some caution in interpreting the stability of the factor structures. In sum, the structure of both the two- and three-factor EFA solutions resembled the bifactor S-1 model with specific peer victimization and reverse-scoring methods factors (Model 4), providing additional support for this model of covariation in childhood maltreatment.

@@

%------------------------------------%
% Discussion: summary
%------------------------------------%

% essentially unidimensional, sans a smaller social factor

%------------------------------------%
% Discussion: previous factor models of maltreatment scales
%------------------------------------%

Our results are consistent with a number of other studies that have found a dominant general factor in measures of childhood maltreatment. Applying a bifactor model to the childhood trauma questionnaire, \cite{spinhoven2014childhood} identified a robust general maltreatment factor that explained away the majority of reliable variance in subscale scores (with the exception of sexual abuse). \cite{meinck2021factor} found similar results when investigating the bifactor structure of the International Child Abuse Screening Tool (ICAST). In a recent investigation of the adverse childhood experience (ACE) questionnaire, \cite{dobson2021latent} found that on average items loaded more strongly onto the general factor than onto their respective specific factors. Collectively these results echo the co-occurrence of types of childhood maltreatment across measurement tools.

>>>>

% override apa6 section headers
% https://tex.stackexchange.com/questions/125537/how-to-modify-subsubsection-header-apa6-cls
\makeatletter
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}
  {\z@}%
  {\b@level@two@skip}{\e@level@two@skip}%
  {\normalfont\normalsize\bfseries}}
\makeatother

@@

\title{The `Maltreatment and Abuse Chronology of Exposure' (MACE) scale is essentially unidimensional}
\shorttitle{Bifactor models of MACE}
\author{Samuel Zorowitz$^1$, Lauri Tuominen$^{2}$}
\affiliation{$^1$Princeton Neuroscience Institute, Princeton University, USA\\$^2$The Royal’s Institute of Mental Health Research, University of Ottawa, Canada}

@@

\begin{document}

\maketitle

@@

Childhood maltreatment is an unfortunately common experience for children across the globe \citep{stoltenborgh2015prevalence}. It is a risk factor for multiple poor outcomes later in life, including the onset of mental illness \citep{macmillan2001childhood, green2010childhood, kessler2010childhood}, challenges in cognitive functioning \citep{kavanaugh2017neurocognitive, r2018common, su2019does}, and poorer physical health \citep{wegman2009meta, widom2012prospective, goodwin2004association}. Childhood maltreatment also predicts worse treatment outcomes for many psychiatric disorders \citep{nanni2012childhood, thomas2019childhood, schuckher2019history}. Thus, the reliable and valid assessment of childhood maltreatment is an important research goal in order to refine our understanding of the link between maltreatment and health outcomes; augment treatment strategies for victims of childhood maltreatment; and ultimately design better interventions to prevent maltreatment in the first place.

There already exist many measures of childhood abuse and maltreatment for researchers to choose from \citep{saini2019systematic}. One recently developed measure, the Maltreatment and Abuse Chronology of Exposure (MACE) scale \citep{teicher2015maltreatment}, has a number of notable advantages. The MACE is a 52-item retrospective self-report questionnaire comprised of 10 short subscales that each assess a unique type of childhood maltreatment. Six of these (e.g. parental verbal abuse, physical abuse, nonverbal emotional abuse, sexual abuse, emotional neglect, physical neglect) overlap with other common measures of abuse; the remaining four assess forms of maltreatment that are less frequently represented in other measures (e.g. peer victimization, witnessing domestic violence). The items in the MACE were carefully selected to measure adverse childhood experiences of increasing severity in order to more finely differentiate people of varying exposure levels. The MACE also measures the timing of maltreatment experiences, which is critical for investigating how the onset and duration of abuse shapes trajectories of development. The MACE provides two measures, an overall severity score (number of maltreatment experiences) and a multiplicity score (number of types of maltreatment experienced), with excellent temporal stability, good convergent validity with other standard measures, and improved predictive validity in predicting psychiatric symptom ratings \citep{teicher2015maltreatment}. For these reasons, the MACE has been listed as among the best measures of childhood maltreatment currently available \citep{saini2019systematic, georgieva2022systematic}.

In recent years, there have been calls for childhood maltreatment research to move away from the cumulative risk framework \citep{evans2013cumulative} --- which focuses only on the number of adverse experiences --- towards a dimensional approach, which emphasizes how distinct types of maltreatment uniquely shape developmental trajectories \citep{mclaughlin2016beyond, belsky2012beyond}. In response, some researchers have begun to use the MACE to form new indices of maltreatment in lieu of or in addition to the severity and multiplicity scores. For example, some have investigated the unique contributions of individual differences on all 10 MACE subscale scores to the development of psychopathology \citep{schalinski2015type, gerke2018childhood, schalinski2019early}. Following a prominent dimensional model of maltreatment \citep{mclaughlin2014childhood}, other groups have made composite indices of threat and deprivation experiences using items in the MACE that respectively assess abuse and neglect experiences \citep{schalinski2018defining, schalinski2019environmental, teicher2018differential}. Though these alternatives maltreatment scores may yield new insight, their psychometric properties have not been extensively investigated.

One critical challenge in measuring childhood maltreatment is the high co-occurrence adverse experiences. That is, children experiencing one form of maltreatment (e.g. physical abuse) are likely to experience one or more other types (e.g. emotional abuse; \citealt{green2010childhood, mclaughlin2012childhood, marques2021risk, mersky2017rethinking}). As a consequence, isolating the unique contribution of one type of maltreatment to the development of some negative outcome is difficult insofar as any one type of adversity is likely confounded with general maltreatment. Concretely, this means that researchers who want to study the effects of particular types of maltreatment using subscale scores or composite indices need to verify those measures reflect a construct mostly independent of overall maltreatment, lest they draw spurious conclusions.

Bifactor models, one type of hierarchical factor models, are well-suited for addressing this issue \citep{bornovalova2020appropriate}. They assume that the covariance structure of a set of items can be accounted for by a general factor, reflecting shared variance across all items, and one or more specific factors, reflecting additional shared variance among subsets of items \citep{Reise2012-ql}. Crucially, bifactor models can be used to calculate model-based reliability statistics, including the omega hierarchical ($\omega_h$) and omega subscale ($\omega_s$) indices, which respectively measure the proportion of variance in total and subscale scores attributable to the general factor and specific factors \citep{reise2013applying, rodriguez2016evaluating}. Applied to the MACE scale, these indices can inform researchers as to whether subscale scores or composite indices for particular types of maltreatment are interpretable as such, or if they instead predominantly reflect overall maltreatment.

% summary of paper paragraph tk

\section{Methods}

@@

The second sample of participants (N=687) were recruited from the Prolific Academic platform (\url{https://www.prolific.co}) to participate in an online experiment in August -- October, 2021. Of these, N=582 were retained for analysis (see Exclusion Criteria below). Participants were eligible for participation if they were 18 years or older and currently resided in the United States or Canada. Participants received monetary compensation for their time (rate: \$12 USD/hr). This study was approved by the Institutional Review Board of the University of Ottawa (\#2021003), and all participants provided informed consent. This sample will hereafter be referred to as the \textit{replication} sample.

@@

The goodness-of-fit indices for each model and sample are summarized in Table \ref{table:diagnostics}. According to the traditional indices, each model provided at least an acceptable fit to the response data. All RMSEA value were <0.06 and the majority of CFI and TLI values were >0.95. The only exception were for the fits of models 1 and 3 to the original sample data, which exhibited only acceptable fits (i.e. CFI/TLI > 0.90). In general, model fits were better for the replication sample data than for the original sample. Furthermore, all of the posterior predictive p-values corresponding to the $\chi^2_{NC}$ discrepancy measure were within the acceptable range. Thus all models were sufficiently able to reproduce the distribution of overall severity scores for each sample.

Next we inspected the $Q_3$ indices for evidence of local dependence. For all models and samples, the maximum observed $Q_3$ index was greater than the 99\% percentile critical value. To understand the extent of local dependence for each model and sample, we visualized the $Q_3$ values for item pairs exceeding the critical value (Figure S3). The unidimensional model fits exhibited many locally dependent item pairs, especially among the peer victimization and reverse-scored items, suggesting possible underfactorization. By comparison, models 2 -- 4 exhibited many fewer locally dependent item pairs. Of these, the reasons for dependence were clear. For example, the dependence between items 17 (``Parents made inappropriate sexual comments or suggestions to your sibling or stepsibling'') and 18 (``Parents touched or fondled your sibling or stepsibling in a sexual way'') present for all models is easily understood as surface local dependence. Regardless, the estimated item discrimination parameters for each model fit seldom exceeded the normal range ($\alpha \leq 4$). That is, the degree of local dependence present in the models is not yielding substantial bias in parameter estimation \citep{edwards2018diagnostic}. Thus, we conclude that all model fits exhibit at least an acceptable fit to the response data.

The results of the model comparison are also summarized in Table \ref{table:diagnostics}. Across samples, models 2 and 4 provided better fits to the data than models 1 and 3. Model 2 provided a numerically better fit to the data than model 4 in the original sample ($\Delta \text{LOO}$ = 7.9, se = 56.3), but the opposite was observed in the replication sample ($\Delta \text{LOO}$ = -24.0, se = 52.0). These differences in LOO-CV values, however, are both within four standard errors of the mean indicating only weak predictive improvements \citep{vehtari2022cv}. Thus the results suggest that models 2 and 4 provide approximately equal fits to the data, which is notable given that model 4 assumes a more parsimonious factor structure of the MACE.

\subsection{Confirmatory item response models}

In the following sections, we interpret the fit of each confirmatory item response model. To ease interpretation, we have reproduced the standardized factor loadings of each model for the original sample in Figure \ref{fig:loadings_original} and for the replication sample in Figure \ref{fig:loadings_online}. The model-based reliability indices for the bifactor models are summarized in Table \ref{table:reliability}.

\subsubsection{Model 1: Unidimensional model of the MACE}

The item discrimination parameters of the unidimensional model were, on average, large for both the original sample (mean $\alpha$ = 1.357, 95\% HDI = 1.295 -- 1.414) and replication sample (mean $\alpha$ = 1.349, 95\% HDI = 1.272 -- 1.422). The standardized factor loadings were correspondingly large in magnitude. In the original sample, factor loadings ranged between 0.279 and 0.864, with an average value of 0.597 (95\% HDI = 0.579 -- 0.614); in the replication sample, factor loadings ranged between 0.359 and 0.814, with an average value of 0.600 (95\% HDI = 0.578 -- 0.621). Most of the items had loadings greater than 0.50 (original: 80.2\%, replication: 80.0\%), and thereby appear to be good indicators of general maltreatment. The rank correlation of factor loadings between samples was $\rho$ = 0.565 (95\% HDI = 0.414 -- 0.711), indicating only moderate agreement on the relative loading of items onto the unitary factor across samples.

To provide one perspective on the dimensionality of the MACE, we next inspected the first six eigenvalues of the polychoric correlation matrix for each sample. In the original sample, these values were 14.62, 3.77, 2.64, 2.11, 1.79, and 1.35; in the replication sample, they are 13.27, 4.43, 2.48, 2.04, 1.81, and 1.56. The first-to-second eigenvalue ratios are 3.84 and 3.02, respectively, just above the 3-to-1 criterion often cited in evaluating whether item response data are essentially unidimensional \citep{embretson2013item}.

In sum, the results provide evidence on average in favor a unidimensional model of the MACE. The factor loadings suggest the presence of a single factor running through the items, as do the eigenvalues of the polychoric correlation matrix. Conversely, the degree of local dependence present among in the unidimensional model fits --- especially among the peer victimization items --- suggest there may be secondary dimensions to the MACE. We turn next to the multidimensional confirmatory models to explore these possibilities.

\subsubsection{Model 2: Bifactor model with 10 specific factors}

Next we inspect the bifactor model with 10 specific factors, one per MACE subscale (model 2). In the original sample, the loadings on the general maltreatment factor ranged between 0.235 and 0.864, with an average value of 0.575 (95\% HDI = 0.556 -- 0.593). In the replication sample, the general factor loadings ranged between 0.299 and 0.827, with an average value of 0.580 (95\% HDI = 0.560 -- 0.601). The relative parameter bias was marginal for both samples (original = 2.1\%; replication = 2.7\%), suggesting little bias in the factor loadings of the unidimensional models due to multidimensionality.

By contrast the standardized loadings on the specific factors were much smaller in magnitude, seldom exceeding loadings on the general factor. In the original dataset, the specific factor loadings ranged 0.014 to 0.797 with an average value of 0.313 (95\% HDI = 0.285 -- 0.341). In the replication dataset, the range was between 0.043 and 0.698, with an average value of 0.343 (95\% HDI = 0.315 -- 0.369). Only items from the peer victimization subscales (i.e. peer verbal \& physical abuse) had specific factor loadings of roughly equal or greater magnitude to those on the general factor.

Given these, we found that the MACE overall severity score had excellent reliability (original: $\omega$ = 0.963; replication: $\omega$ = 0.965). Further, the $\omega_h$ value for both samples was 0.924 indicating that 96.0\% and 95.8\% of the reliable variance in the overall severity score is attributable to the general maltreatment factor. (This is not in itself so surprising given the large number of subscales and relative paucity of items per subscale.) Notably, the ECV values were 0.716 and 0.697 for the original and replication samples, respectively. In conjunction with a large PUC value of 0.912, these results provide clear evidence for a dominant general maltreatment factor in the MACE.

The reliability of the MACE subscale scores were more variable in comparison. The $\omega$ values ranged between 0.590 and 0.909 in the original sample, and between 0.595 and 0.893 in the replication sample, with smaller values for subscales with fewer items. Crucially, the $\omega_s$ values were mostly small for both the original sample (mean $\omega_s$ = 0.181, range = 0.027 -- 0.621) and replication sample (mean $\omega_s$ = 0.205, range = 0.037 -- 0.565). That is, the MACE subscale scores contain almost no reliable variance independent of the general maltreatment factor. Therefore they should not be interpreted as reflecting specific forms of maltreatment. The only exception was for the peer verbal abuse subscale, where more than half the reliable variance in subscale scores (original = 70.7\%; replication = 66.2\%) was independent of general maltreatment.

The results of the bifactor model were unequivocal. Responses on the MACE are characterized by a dominant general maltreatment factor with strong loadings from the majority of items. In stark contrast, the specific factors corresponding to each MACE subscale are weak and characterized mostly by small factor loadings. Whereas the MACE overall severity score is reliable and a valid measure of general maltreatment, the MACE subscale scores are less reliable and cannot be meaningfully interpreted as uniquely reflecting exposure to specific types of abuse. The question remains, however, if there are other valid composite scores to be derived from the MACE in addition to the overall severity score.

\subsubsection{Model 3: Bifactor S-1 model with one specific factor (deprivation)}

We turn our attention next to the threat-deprivation bifactor S-1 model (model 3). The loadings on the general factor largely resembled those of the bifactor model. In the original sample, the loadings on the general maltreatment factor ranged between 0.150 and 0.864, with an average value of 0.577 (95\% HDI = 0.558 -- 0.594). In the replication sample, the general factor loadings ranged between 0.395 and 0.831, with an average value of 0.580 (95\% HDI = 0.558 -- 0.601).

The loadings on the specific (neglect) factor were smaller on average. In the original sample, the average specific loading was 0.410 (95\% HDI = 0.373 -- 0.454); in the replication sample, it was 0.454 (95\% HDI = 0.409 -- 0.499). The specific loadings exhibited a notable pattern where only the six reverse-scored items exhibited moderate loadings; the remaining four item exhibited negligible factor loadings. In other words, the ``neglect'' factor appears instead to reflect a reverse-scoring methods factor.

As observed with the bifactor model, the overall severity score under the threat- deprivation bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.958; replication: $\omega$ = 0.959). Unsurprisingly, the $\omega_h$ values indicated the majority of reliable variance in severity score is attributable to the general factor (original: $\omega_h$ = 0.928; replication: $\omega_h$ = 0.922). The reliability of the ``neglect'' composite score was similarly good (original: $\omega$ = 0.877; replication: $\omega$ = 0.921). The $\omega_s$ values, however, indicated the composite score contained less than half reliable variance independent of the general factor (original: $\omega_h$ = 0.384; replication: $\omega_h$ = 0.375).

In summary then, the results do not support the use of composite threat and deprivation scores as derived from the abuse and neglect items on the MACE. The minority of reliable variance unique from general maltreatment prevents the deprivation score from being interpreted as such. Furthermore, the items comprising deprivation composite index seem to be confounded by a reverse-scoring methods factor.

\subsubsection{Model 4: Bifactor S-1 with two specific factors (peer victimization, reverse-scoring)}

Finally, we inspect the bifactor S-1 model with two specific factors measuring peer victimization and reverse-scoring (model 4). In the original sample, the loadings on the general maltreatment factor ranged between 0.158 and 0.864, with an average value of 0.563 (95\% HDI = 0.545 -- 0.580). In the replication sample, the general factor loadings ranged between 0.298 and 0.852, with an average value of 0.572 (95\% HDI = 0.550 -- 0.593). Interestingly, the loadings on the specific factors were of approximately equal magnitude. In the original dataset, the specific factor loadings ranged 0.327 to 0.789 with an average value of 0.550 (95\% HDI = 0.522 -- 0.576). In the replication dataset, the range was between 0.183 and 0.684, with an average value of 0.526 (95\% HDI = 0.492 -- 0.556).

As with the other bifactor models, the overall severity score under this bifactor S-1 model exhibited excellent reliability (original: $\omega$ = 0.961; replication: $\omega$ = 0.961). Again the majority of reliable variance in severity score was attributable to the general factor. The peer victimization composite score exhibited good reliability across samples (original: $\omega$ = 0.889; replication: $\omega$ = 0.868). Notably, the corresponding $\omega_s$ values were 0.569 for the original sample and 0.493 for the replication sample. That is, slightly more than half of the variance in peer victimization scores (original: 64.0\%; replication: 56.8\%) is unique from the general maltreatment factor. These results suggest there exists a secondary factor in the MACE, peer victimization, independent of general maltreatment (albeit marginally so).

Briefly, we note that the composite scores formed by the reverse-scored items was also reliable (original: $\omega$ = 0.839; replication: $\omega$ = 0.915). Moreover, the corresponding $omega_s$ values were similar in magnitude to those observed for the peer victimization scores (original: $\omega_s$ = 0.584; replication: $\omega_s$ = 0.498). As such, these scores are also have majority reliable variance independent of general maltreatment. It is difficult to interpret these scores, however, as it is unclear if they primarily reflect neglect experiences, a wording effect, or a scoring-induced methods artifact.

@@

The standardized factor loadings for the two- and three-factor solutions generated by the cf-quartimax rotation are summarized in Figure \ref{fig:efa}. (The corresponding factor loadings produced by the geomin rotation were similar and are presented in Figure S4.) In the two-factor solution for the original sample, there was a primary adult maltreatment factor and a secondary peer victimization factor. In contrast, the replication sample produced a primary adult/peer maltreatment factor with a secondary reverse-scoring factor.

The three-factor solution for each sample were similar. The analysis produced a primary adult maltreatment factor with secondary peer victimization and reverse-scoring factors. (To note items concerning sexual abuse loaded more strongly on the peer victimization scale in the replication sample, warranting some caution in interpreting the replicability of factor structures.) With the oblique rotation, we observed only modest inter-factor correlations (original: mean $r$ = 0.371; replication: mean $r$ = 0.269). To conclude, the structure of both the two- and three-factor exploratory factor solutions resembled the structure of the bifactor S-1 model with two specific factors (model 4), providing further support for this model of responses on the MACE.

@@

The MACE is an increasingly popular retrospective self-report measure of childhood maltreatment. Despite this, there has been little investigation into its factor structure \citep{saini2019systematic}, raising questions about the validity and interpretability of MACE subscale scores and composite indices. Here we employed a series of confirmatory and exploratory item factor models to investigate the dimensionality of the MACE. Simply put, we found repeated evidence across two independent samples that the MACE is essentially unidimensional. First, one-factor (unidimensional) models of the MACE provided adequate (albeit worse) fits to both datasets. Second, the omega hierarchical ($\omega_h$) for the full bifactor model was 0.924, indicating that the overwhelming majority of variance  in MACE severity scores is attributable to general maltreatment and only trivially affected by multidimensionality of the specific factors. Third, the explained common variance was, on average, approximately 0.70, suggesting that unidimensional models of the MACE are not expected to suffer from seriously biased parameter estimates.

As a corollary, we found little evidence for the reliability of MACE subscale scores or composite indices. Indeed, the omega subscale ($\omega_s$) values indicated that there was little reliable variance in almost all of the MACE subscale scores after accounting for the general factor. That is, MACE subscale scores cannot be interpreted as such but should instead be understood as predominantly reflecting overall maltreatment. The same can be said of the deprivation composite index, which had little reliable variance unique from the general factor and is seemingly contaminated by a reverse-scoring methods factor. Notably, we did find weak evidence for a secondary peer victimization factor in the MACE, whose composite scores possess a small majority of unique variance independent of general maltreatment. The existence of this peer victimization factor was supposed by exploratory factor modeling. Thus, it may be possible to derive two measures from the MACE, maltreatment by parents or other adults and maltreatment by peers, but further research is required to support this conclusion.
